SunPy - Python for Solar Physics

### 2014-01-29 3-3-HELIO.tex

The HELiophysics Integrated Observatory (\href{http://helio-vo.eu}{HELIO}) has 
compiled a list of web services which allows scientists to query and 
discover data throughout the heliosphere, from solar and magnetospheric data to planetary and 
inter-planetary data \citep{dps2012}.
HELIO is built with a Service-Oriented Architecture, 
i.e., its capabilities are divided into a number of tasks that are 
implemented as separate services. 
\href{http://helio-vo.eu}{HELIO} is made up of nine different public services, 
which allows scientists to search different catalogues of registered events, 
solar features, data from instruments in the heliosphere and other information 
such as planetary or spacecraft position in time. 
Additionally, \href{http://helio-vo.eu}{HELIO} provides a service that uses a 
propagation model to link the data in different points of the solar system by 
its original nature (e.g., Earth auroras are a signature of magnetic 
field disturbances produced few days before on the Sun).
In addition to the primary, web-based interface to 
\href{http://helio-vo.eu}{HELIO}, the services are available by API.

SunPy's \texttt{hec} module provides an interface to the
\textit{HELIO Event Catalogue} (HEC) service. This module was developed as
part of a Google Summer of Code (GSOC) project in 2013.
The HEC service currently provides access to 84 catalogues from different
sources.
As with all of the HELIO services, the HEC service provides results in VOTable 
data format (defined by IVOA \citep{ochsenbein_ivoa_2011}). 
%schriste - a short descriptions of what the hec (get it?) a vot table is needed here
The \texttt{hec}
module parses this output using the \texttt{astropy.io.votable} package.
This format has the advantage of containing metadata with information like
data provenance and the performed query.

@@

# Query all the catalogues that comes from cactus
>>> results = [hc.time_query(tstart, tend, event) 
...            for event in catalogues_event if 'cactus' in event]
>>> for cat in results:
...     print "{cat} has {nres} results".format(cat = cat.ID, \
...                                             nres = len(cat.array))
__helio_hec-cactus_stereoa_cme has 4 results
__helio_hec-cactus_stereob_cme has 3 results
__helio_hec-cactus_soho_cme has 7 results
\end{minted}
\caption{Example of querying the HEC service to multiple cme
catalogues, in this case the ones detected by \href{http://sidc.oma.be/cactus/}{CACTus}.}
\label{code:helio}
\end{listing}
% RJH: CACTus should be defined and a reference provided in the reference section.

In the future, additional HELIO services will be supported in the future by SunPy.
% RJH: This sentence is an orphan and should have a home constructed for it.
% schriste - it also contains too much future sauce

>>>>

The HELiophysics Integrated Observatory (\href{http://helio-vo.eu}{HELIO}) has 
compiled a list of web services which allows scientists to query and 
discover data throughout the heliosphere, from solar and magnetospheric data to planetary and 
inter-planetary data \citep{dps2012}.
HELIO is built with a Service-Oriented Architecture, 
i.e., its capabilities are divided into a number of tasks that are 
implemented as separate services. 
\href{http://helio-vo.eu}{HELIO} is made up of nine different public services, 
which allows scientists to search different catalogues of registered events, 
solar features, data from instruments in the heliosphere and other information 
such as planetary or spacecraft position in time. 
Additionally, HELIO provides a service that uses a 
propagation model to link the data in different points of the solar system by 
its original nature (e.g., Earth auroras are a signature of magnetic 
field disturbances produced few days before on the Sun).
In addition to the primary, \href{http://hfe.helio-vo.eu}{web-based interface to 
HELIO}, the services are available by API.

SunPy's \texttt{hec} module provides an interface to the
\textit{HELIO Event Catalogue} (HEC) service. This module was developed as
part of a Google Summer of Code (GSOC) project in 2013.
The HEC service currently provides access to 84 catalogues from different
sources.
As with all of the HELIO services, the HEC service provides results in VOTable 
data format (defined by IVOA \citep{ochsenbein_ivoa_2011}).
%schriste - a short descriptions of what the hec (get it?) a vot table is needed here
%dps - a votable format description? or what HEC gets from it?
The \texttt{hec}
module parses this output using the \texttt{astropy.io.votable} package.
This format has the advantage of containing metadata with information like
data provenance and the performed query.

@@

# Query all the catalogues that comes from cactus
>>> results = [hc.time_query(tstart, tend, event) 
...            for event in catalogues_event if 'cactus' in event]
>>> for cat in results:
...     print "{cat} has {nres} results".format(cat = cat.ID, \
...                                             nres = len(cat.array))
__helio_hec-cactus_stereoa_cme has 4 results
__helio_hec-cactus_stereob_cme has 3 results
__helio_hec-cactus_soho_cme has 7 results
\end{minted}
\caption{Example of querying the HEC service to multiple cme
catalogues, in this case the ones detected automatically 
by \href{http://sidc.oma.be/cactus/}{CACTus} \citep{robbrecht_automated_2009}.}
\label{code:helio}
\end{listing}

In the future, additional HELIO services will be supported by SunPy, aiming to a 
seamlessly interaction between the other services and tools available (e.g., 
\texttt{hek}, \texttt{Map}).  
For example, solar features available in the HELIO Feature Catalogue could be
overplot on maps.
% RJH: This sentence is an orphan and should have a home constructed for it.
% schriste - it also contains too much future sauce
% better? or this should be in the future section?

### 2014-01-29 3-4-Helioviewer.tex

SunPy provides the ability to download images hosted by the
\href{http://helioviewer.org}{Helioviewer Project}.  
The aim of the Helioviewer Project is to enable
the exploration of solar and heliospheric data from multiple data
sources (such as instrumentation and feature/event catalogues) via
easy-to-use visual interfaces.  The primary data provided by the
Helioviewer Project are images in the JPEG2000 format (for further
details on the Helioviewer Project see \cite{muller2009}). The
JPEG2000 files are typically highly compressed compared to the source
FITS files they are generated from, and so can be a used as a shortcut
to visualise large amounts of data from multiple sources.  SunPy is
also used in Helioviewer production servers to manage the download and
ingestion of JPEG2000 files from remote servers.

@@

%Stuart: Do we need this listing?? The same points can be made with 14
\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy.net.helioviewer import HelioviewerClient
>>> hv = HelioviewerClient()
>>> hv.download_png('2099/01/01', 4.8, "[SDO,AIA,AIA,304,1,100]",
...                 x0=0, y0=0, width=512, height=512)
\end{minted}
\centering\includegraphics[width=0.5\columnwidth]{helioviewer_latest_aia_304}
\caption{Acquisition of a PNG file showing the latest AIA 304\AA\ image 
available at \url{www.helioviewer.org}.}
\label{code:hv:downloadlatestpng}
\end{listing}

@@

It should be noted that the filename of the returned file has the date
and time of the requested time, which is not the time of the observation shown 
in the PNG image.  This is because the Helioviewer API request finds images 
closest to the requested time. However, since the user may ask for images from 
multiple sources, and each of them may have a different observation time, it is
unclear which time is the most appropriate to associate with the resultant 
image.  The Helioviewer API therefore does not select between any of the image 
observation times, but instead returns an image filename based on the request 
time, resulting in predictable filenames.

>>>>

SunPy provides the ability to download images hosted by the
\href{http://helioviewer.org}{Helioviewer Project}.  
The aim of the Helioviewer Project is to enable
the exploration of solar and heliospheric data from multiple data
sources (such as instrumentation and feature/event catalogues) via
easy-to-use visual interfaces.  The primary interface to Helioviewer is a
google-map-like website which allows users to browse images of the Sun
taken by a variety of instruments. In order to manage all of this data, the Helioviewer
project maintains archives of all of its data as JPEG2000 format (for further
details on the Helioviewer Project see \citep{muller2009}). The
JPEG2000 files are typically highly compressed compared to the source
FITS files they are generated from, and so can be a used as a shortcut
to visualise large amounts of data from multiple sources.  SunPy is
also used in Helioviewer production servers to manage the download and
ingestion of JPEG2000 files from remote servers.

@@

%Stuart: Do we need this listing?? The same points can be made with 14
% schriste - i agree with stuart here, or maybe just don't show the png file
\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy.net.helioviewer import HelioviewerClient
>>> hv = HelioviewerClient()
>>> hv.download_png('2099/01/01', 4.8, "[SDO,AIA,AIA,304,1,100]",
...                 x0=0, y0=0, width=512, height=512)
\end{minted}
\centering\includegraphics[width=0.5\columnwidth]{helioviewer_latest_aia_304}
\caption{Acquisition of a PNG file showing the latest AIA 304\AA\ image 
available at \url{www.helioviewer.org}.}
\label{code:hv:downloadlatestpng}
\end{listing}

@@

It should be noted that the filename of the returned file has the date
and time of the requested time, which is not the time of the observation shown 
in the PNG image.  This is because the Helioviewer API request finds images 
closest to the requested time. However, since the user may ask for images from 
multiple sources, and each of them may have a different observation time, it is
unclear which time is the most appropriate to associate with the resultant 
image.  The Helioviewer API therefore does not select between any of the image 
observation times, but instead returns an image filename based on the request 
time, resulting in predictable filenames.
%this seems like a lot of text for a simple issue, i would recommend removing it
